---
title: 'Exercise #2'
author: "Raffaello Lastrico"
date: "date of submission"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# more internal settings can go here
# Consider help pages like:
# https://rmarkdown.rstudio.com/lesson-1.html
# https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf

```

### Data
Analysis based on this timeseries:
<https://github.com/data-hydenv/data/blob/master/hobo/2022/10_minute/10347366.csv>

### Loaded packages
tidyverse, lubridate, hms, zoo, readr

```{r libraries, include=FALSE}
library(tidyverse)
library(lubridate)
library(zoo)
library(readr)
library(hms)
```

### Preprocessing / R functions

```{r preprocessing_or_functions, message=FALSE, warning=FALSE}
QCP4 <- function(x, threshold){
    if(max(x, na.rm = T) >= threshold){
        return(1)
      }
    else{
        return(0)
      }
}
```

## 1. Quality control procedures (4 QCPs)

```{r, eval = TRUE, message=F}
# Load data from Github
data <- read_csv(file = "https://raw.githubusercontent.com/data-hydenv/data/master/hobo/2022/10_minute/10347366.csv")
head(data)
```

### 1.1 Measurement range (Plausible values)

**HOBO Specifications:**
Temperature: -20° to 70° C  -  Light intensity: 0 to 320000 lux

```{r qcp1, eval = TRUE}
# Flag values outside the ranges with 1, those within with 0 
data <- data %>% 
  mutate(QCP1 = case_when(temp <= 70 & temp >= -20 & lux <= 320000 & lux >= 0 ~ 0,
                          TRUE ~ 1)
        )

sum(data$QCP1)

```

**Question**: How many data points are outside the measurement range?

**Answer**: There are 0 data points outside the measurement range of both temperature and light intensity.

### 1.2 Plausible rate of change

```{r qcp2, eval=T}
# Calculate diference in temperature und flag those > 1° C
data <- data %>% 
  #head() %>% 
  mutate(var1 = temp - lag(x = temp), 
         QCP2 = if_else(abs(var1) > 1, true = 1, false = 0)
         ) %>% 
  select(-var1)

# Summarise time and count information for QCP2 classification
data %>% 
  mutate(daytime = strftime(dttm, format="%H:%M:%S")) %>% 
  group_by(QCP2) %>% 
  summarise(minTime = min(daytime), maxTime = max(daytime), count = length(daytime))
```

**Question**: Describe shortly how many data points failed during this QCP and discuss whether there is a certain daytime pattern of failure or not?

**Answer**: 232 data points failed this quality criterion. 3799 passed. 1 value was assigned NA, due to it being the first point. The measurements that failed all occurred between 7 am and 6:10 pm. This might suggest that larger temperature increases or drops are influenced by sunlight. 

### 1.3 Minimum variability (Persistence)

```{r qcp3}
# code for qcp3 here
data <- data %>%
  mutate(QCP3 = case_when(temp == lag(temp, n = 1) & temp == lag(temp, n = 2) & 
                          temp == lag(temp, n = 3) & temp == lag(temp, n = 4) &
                          temp == lag(temp, n = 5) ~ 1,
                          TRUE ~ 0)
        )

sum(data$QCP3, na.rm = T)

# Question: with this solution the first values of the series are not marked NA in QCP3. Is that a problem? 
```

**Task**: Code in this section should analyse the persistence.

*Results:* The time series contains 47 periods of continuos 10 minute intervals with the same value.

### 1.4 Light intensity

```{r qcp4}
# classify data points by light intensity and period of the day (day/night)
data <- data %>% 
  mutate(SIC = case_when(
                lux >= 0 & lux <= 10 ~ "0_night",
                lux >= 10 & lux <= 500 ~ "1_sunrise_sunset",
                lux >= 500 & lux <= 2000 ~ "2_overcast_full",
                lux >= 2000 & lux <= 15000 ~ "3_overcast_light",
                lux >= 15000 & lux <= 20000 ~ "4_clear_sky",
                lux >= 20000 & lux <= 50000 ~ "5_sunshine",
                lux >= 50000 ~ "6_sunshine_bright",
                TRUE ~ "NA"
               ),
         time = as_hms(dttm),
         period = if_else(time >= as_hms("6:00:00") & time <= as_hms("17:59:59"), true = "day", false = "night")
         ) %>% 
  relocate(time, period, .after = dttm) %>% 
  relocate(SIC, .after = lux)

# Count observations within light intensity classes
data %>% count(SIC)

# Implement QCP 4 criterion: flag day time observations >20000 lux (i +- 1); >50000 (i +- 50000)
data <- data %>%
  mutate(QCP4_1 = if_else(period == "day", true = rollapply(lux, width = 3, FUN = QCP4, threshold = 20000, fill = NA, align = "center"), false = 0), 
         QCP4_2 = if_else(period == "day", true = rollapply(lux, width = 7, FUN = QCP4, threshold = 50000, fill = NA, align = "center"), false = 0),
         QCP4 = if_else(QCP4_1 + QCP4_2 > 0, true = 1, false = 0)) %>% 
  select(-QCP4_1, -QCP4_2)

# Visualize, when the QCP4 criterion fails (which times of day)
data %>% 
  filter(period == "day") %>% 
  group_by(hour = hour(dttm)) %>% 
  summarise(QCP4_Count = sum(QCP4))%>%
  ggplot(aes(x = hour, y = QCP4_Count))+
    geom_col()+
    scale_x_continuous(breaks = seq(6,18, by = 1))+
    xlab("Time")+
    ylab("Count of QCP4 fails")

# Calculate total number of fails
data %>% 
  count(QCP4)

```

**Task**: Discuss shortly how often and when during daytime the QCP4 flags bad data. Elaborate on some reasons for your results.

**Answer**: The QCP4 criterion fails mostly around noon (about 150 observations), but never before 9 am or after 3 pm. Over the whole time span 272 observations out of 4032 fail. 

## 2. Synthesis

```{r synthesis}
# compare combinations of failures
data %>%
  count(QCP1,QCP2,QCP3,QCP4)

```

**Task**: Present a table or graph to show how many data points fail during the four specific QCPs. Discuss shortly the reasons for failure and compare the different QCPs against each other.

**Answer**: The above output shows, that 3662 10-min observations pass all 4 criteria. QCP1 was always passed, which was to be expected as the sensor only recorded values within its measurement range. QCP2 failed 232 times. QCP3 only failed 47 times. QCP4 failed 272 times, which is to be explained by the sensors position on a south facing wall.

## 3. Results

### 3.1 Result (Flagging system: 10-minutes-values)

```{r res1}
# create cumulated flags
data <- data %>% 
            mutate(QCP_cum = if_else(QCP1 + QCP2 + QCP3 + QCP4 > 0, true = 1, false = 0))

# select required variables
qc_df <- data %>% 
            select(id, dttm, temp, lux, QCP1:QCP_cum)

head(qc_df,100)
```

**Task**: At the end of the code section above you should generate one! tibble or data.frame named `qc_df` with all time information, all data points (temperature and lux) and your outcomes of the different QCPs.

### 3.2 Result (Aggregate to hourly series)

```{r res2}
# create qrouping variable
qc_df <- qc_df %>% 
            mutate(day_hour = ymd_h(format(dttm, format = "%Y-%m-%d %H")))

# aggregate qc_df to hourly values (condition: not more than 1 QCP_cum failure per hour)
hobo_hourly <- qc_df %>% 
            group_by(day_hour) %>% 
            summarise(temp = if_else(sum(QCP_cum, na.rm = T) <= 1, true = round(mean(temp), 4), false = NULL)) %>% 
            rename(dttm = day_hour)

head(hobo_hourly, 50)

```

**Task**: At the end of the code section above you should generate one! tibble or data.frame named `hobo_hourly` with averaged temperature values per hour or NA values (if the hour is flagged as bad data). See exercise description for more details.

-   First column: YYYY-DD-MM HH:MM:SS

-   Second column: Temperature values (4 digits), NA values possible
